---
title: "PML Project"
output: html_document
---

## Introduction 
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.

## Getting data
The data for this project come from this [source](http://groupware.les.inf.puc-rio.br/har).
The training data are available [here](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv) and the test data [here](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv)

```{r}
downloadcsv <- function(url, nastrings) {
    temp <- tempfile()
    download.file(url, temp, method = "curl")
    data <- read.csv(temp, na.strings = nastrings)
    unlink(temp)
    return(data)
}
#train.url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
#train <- downloadcsv(train.url, c("", "NA", "#DIV/0!"))
train <- read.csv("pml-training.csv")
#test.url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
#test <- downloadcsv(test.url, c("", "NA", "#DIV/0!"))
test <- read.csv("pml-testing.csv")
```
the train data contains 19622 samples, each with 160 features
```{r}
dim(train)
```

## Data Preprocessing
We separate our training data into a training set and a cross-validation set.
```{r}
library(caret)
set.seed(1234)
inTrain <- createDataPartition(y = train$classe, p=0.7, list=F)
training <- train[inTrain,]
cv <- train[-inTrain,]
```

## Feature Engineering
We exclude near-zero variance features and features with too many missings, also excludes description features
```{r}
nzv <- nearZeroVar(training)
training <- training[, -nzv]

count.missing <- sapply(training, function(x) sum(is.na(x) | x==""))
col.remove <- names(count.missing[count.missing > 0.5 * length(training$classe)])
col.desc <- c("user_name", "X", "raw_timestamp_part_1", "raw_timestamp_part_2", 
    "cvtd_timestamp", "new_window", "num_window")
excluded <- c(col.desc, col.remove)
training <- training[, !names(training) %in% excluded]
```

## Model Training
```{r}
library(randomForest)
rf <- randomForest(classe~., data=training, importance = T)
```

## Cross Validation
### Training Set Accuracy
```{r}
print(rf)
```
Our model performs excellent against the training set,with OOB error rate 0.47%.

### Validation Set Accuracy
```{r}
cv.pred <- predict(rf, cv)
confusionMatrix(cv.pred, cv$classe)
```
The cross validation accuracy is 0.995, seems rather good. So the out-of-sample error is 0.5%.

## Prediction on Test Set

```{r}
test.pred <- predict(rf, test)
test.pred
```

We then write results to file as instructed.
```{r}
ans <- as.vector(test.pred)
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(ans)
```

